Trajectory model of passive kelp dispersal

### 2020-08-18 thesis_chapter_3.Rmd
error in citations fixed. outputs updated.

---
title: 'Trajectory model of passive kelp dispersal'
author: "Ross Coppin"
output:
  pdf_document:
    fig_caption: yes
  html_document:
    df_print: paged
  word_document:
    reference_docx: markdown/thesis_template.docx
bibliography: references.bib
---

@@

Many past authors have also taken various approaches to estimate not only the raft-time of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory. Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known; although the role of wind has been recognised as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterrey Bay. The results showed that kelp rafts with little surface area exposed to the wind were largely driven by a combination of wind and wind waves, however the relative importance of wind and wind waves was not clear. In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter. Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @Fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft. The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate; while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].

@@

To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account. Past research by @plourde1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model. However, this work does not consider biological material such as macroalgae. Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates. Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulae to calculate drag based on shape. Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'. A spherecity factor is the ratio of the surface area of a sphere having the volume as the irregular shaped particle (s) to the actual surface area of the particle (s_d) such that $$\psi=s/s_d$$. The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work. A table of various shapes and the associated spherecity factor is provided in the appendix.

>>>>

---
title: 'Trajectory model of passive kelp dispersal'
author: "Ross Coppin"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
  word_document:
    reference_docx: markdown/thesis_template.docx
bibliography: references.bib
---

@@

Many past authors have also taken various approaches to estimate not only the raft-time of macroalgal rafts but also the oceanographic influencers of macroalgae trajectory. Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005, @thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influencer, the relative importance of wind versus surface current is still not known; although the role of wind has been recognised as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterrey Bay. The results showed that kelp rafts with little surface area exposed to the wind were largely driven by a combination of wind and wind waves, however the relative importance of wind and wind waves was not clear. In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter. Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft. The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate; while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].

@@

To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account. Past research by @allen1999 and @breivik2011 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model. However, this work does not consider biological material such as macroalgae. Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates. Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulae to calculate drag based on shape. Chemical engineers in the field of fluid dynamics which come across similar problems when including drag use an approach known as a 'spherecity factor'. A spherecity factor is the ratio of the surface area of a sphere having the volume as the irregular shaped particle (s) to the actual surface area of the particle (s_d) such that $$\psi=s/s_d$$. The values are dependent on the shape of the object and a range of values have been determined based on numerous experimental work. A table of various shapes and the associated spherecity factor is provided in the appendix.
